Experiences with planning for natural language generation

### 2008-06-20 trunk/Papers/experiences/discussion.tex
small change to introduction; changed to discussion section

The positive conclusion we can draw from the experiments reported
above is that for the most part, modern planners are pretty good at
controlling the search for the NLG-related inputs that we looked at.
This is particularly true for FF in the sentence generation domain,
which generates nontrivial 14-word sentences in about two seconds.
There is still room for improvement (a special-purpose algorithm for
referring expressions could generate ten of these 14 words in a few
milliseconds), and the search becomes slower -- for obvious reasons --
in the almost completely unconstrained world of Experiment 2, but
search is not the main efficiency bottleneck.

The most restrictive bottleneck in the two domains we have looked at
in this paper is the initial grounding step that almost all modern
planners perform.  Our conclusion from the experiments reported above
is that this is a good strategy in domains where the typical plan
length dominates the typical universe size, i.e.\ where the initial
investment into grounding pays off in saved unifications during the
search.  However, as we have shown, there are natural planning domains
in which relatively short plans must be computed in large universes;
it is not unrealistic for the universe in the sentence generation
domain to consist of thousands of domain individuals and tens of
thousands of actions, some of which take three domain individuals as
parameters.  The common strategy of splitting the universe over
several types of individuals will not be very effective here.

>>>>

The positive conclusion we can draw from the experiments reported above is
that modern planners are fairly good at controlling the search for many of
the moderately-sized NLG problems we looked at. This is particularly true
for FF in the sentence generation domain, which generates nontrivial
14-word sentences in about two seconds. There is still room for
improvement, however. For instance, by comparison a special-purpose
algorithm for referring expressions could generate ten of these 14-word
sentences in a few milliseconds. Although search efficiency is not the main
performance concern in these domains, it nevertheless becomes a factor in
scenarios like Experiment~2, where the world is almost completely
unconstrained and (unsurprisingly) the search becomes much slower.

%The positive conclusion we can draw from the experiments reported
%above is that for the most part, modern planners are pretty good at
%controlling the search for the NLG-related inputs that we looked at.
%This is particularly true for FF in the sentence generation domain,
%which generates nontrivial 14-word sentences in about two seconds.
%There is still room for improvement (a special-purpose algorithm for
%referring expressions could generate ten of these 14 words in a few
%milliseconds), and the search becomes slower -- for obvious reasons --
%in the almost completely unconstrained world of Experiment 2, but
%search is not the main efficiency bottleneck.

The most restrictive bottleneck in the two domains we investigated is the
initial grounding step that many modern planners perform. While this may be
a good strategy for traditional planning benchmarks and IPC domains---where
plan length often dominates universe size, and an initial investment into
grounding pays off in saved instantiations during the search---this
approach is less efective in the domains we've considered. As our
experiements have shown, there are natural planning domains in which
relatively short plans must be computed in large universes. For instance,
in the case of the sentence generation domain, it is not unrealistic for
the universe to consist of thousands of domain individuals and tens of
thousands of actions, some of which require three domain individuals as
parameters. The common strategy of splitting the universe over several
types of individuals will not be very effective here.

%The most restrictive bottleneck in the two domains we have looked at
%in this paper is the initial grounding step that almost all modern
%planners perform.  Our conclusion from the experiments reported above
%is that this is a good strategy in domains where the typical plan
%length dominates the typical universe size, i.e.\ where the initial
%investment into grounding pays off in saved unifications during the
%search.  However, as we have shown, there are natural planning domains
%in which relatively short plans must be computed in large universes;
%it is not unrealistic for the universe in the sentence generation
%domain to consist of thousands of domain individuals and tens of
%thousands of actions, some of which take three domain individuals as
%parameters.  The common strategy of splitting the universe over
%several types of individuals will not be very effective here.

### 2008-06-20 trunk/Papers/experiences/introduction.tex
small change to introduction; changed to discussion section

We will then discuss some of our experiences using off-the-shelf planners
in these domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN \cite{hsu06:_new_featur_in_sgplan_for} on
varying-sized instances of our planning domains. We also compare these
results against an ad-hoc Java implementation of GraphPlan.  Our findings
are mixed. On the one hand, it turns out that modern planners handle some
of the search problems that arise in NLG quite easily (although large
problem instances remain a challenge). On the other hand, these same
planners spend tremendous amounts of time on preprocessing. For instance,
FF spends 90\% of its runtime in the sentence generation domain on
grounding out literals and actions, and ends up slower than our version of
GraphPlan that avoids grounding. For domains like ours, which are dominated
by the number of actions and the universe size, rather than the
combinatorics of the search problem, this observation suggests that the
overall runtime of a planner could be improved by grounding more
selectively. We therefore offer these domains as challenges for the
planning community.

>>>>

We will then discuss some of our experiences using off-the-shelf planners
in these domains. In particular, we explore the efficiency of FF
\cite{HoffmannNebel01} and SGPLAN
\cite{hsu06:_new_featur_in_sgplan_for}---planners that have been successful
on traditional benchmarks, and in the International Planning
Competition (IPC)---on varying-sized instances of our planning domains. We
also compare these results against an ad-hoc Java implementation of
GraphPlan.  Our findings are mixed. On the one hand, it turns out that
modern planners handle some of the search problems that arise in NLG quite
easily (although large problem instances remain a challenge). On the other
hand, these same planners spend tremendous amounts of time on
preprocessing. For instance, FF spends 90\% of its runtime in the sentence
generation domain on grounding out literals and actions, and ends up slower
than our version of GraphPlan that avoids grounding. For domains like ours,
which are dominated by the number of actions and the universe size, rather
than the combinatorics of the search problem, this observation suggests
that the overall runtime of a planner could be improved by grounding more
selectively. We therefore offer these domains as challenges for the
planning community.
